Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-06-02 paper/paper.tex
removed typo
datarelease

\subsection{prior}
Using \xd\ and the MML method described in the methods section, we generate the prior. This is shown in Figure \ref{fig:prior}.
This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution. The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from. The main sequence and red giant branch are much tighter, as well as the red clump. We'll come back to more features in the data below, especially those seen in the posterior distribution.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{prior_ngauss128.png}
\caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data. The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours. The right panel shows the 1-$\sigma$ contours of the individual components. The latter are Gaussian in the transformed magnitude space to make the XD inference tractable. Thus, they appear non-Gaussian in the physical color--magnitude space.}
\label{fig:prior}
\end{figure}

@@

%\begin{figure}
%\centering
%  \includegraphics[width=\textwidth]{priorNgaussComparison.png}
%\caption{Prior: the XD generated mixture of gaussians we use as our prior. To determine the number of gaussians we used cross validation, using 70\% of our %sample as a training set and 30\% as the test set. Increasing the number of gaussians we used continued to increase the likelihood of the test set of data, %showing that we are no over fitting. Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the %features of the CMD.}
%\label{fig:priorCompare}
%\end{figure}

>>>>

\subsection{XD-generated Prior}
Using \xd\ and the MML method described in the methods section, we generate the prior. This is shown in Figure \ref{fig:prior}.
This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution. The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from. The main sequence and red giant branch are much tighter, as well as the red clump. We'll come back to more features in the data below, especially those seen in the posterior distribution.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{prior_ngauss128.png}
\caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data. The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours. The right panel shows the 1-$\sigma$ contours of the individual components. The latter are Gaussian in the transformed magnitude space to make the XD inference tractable. Thus, they appear non-Gaussian in the physical color--magnitude space.}
\label{fig:prior}
\end{figure}

@@

\subsection{Data products}

The method produces a mixture-of-Gaussian parallax posterior pdf for every
star in the intersection of \tgas, \tmass, and the dust map.
Associated with this paper is an electronic table containing our results, and everything
that is necessary to use them responsibly.
It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
photometry.
It contains posterior expectations and variances for the stellar parallax.
It also contains the full posterior pdf for each star, in the form of a set of pdf values,
evaluated on a common grid in parallax.
Because all of the data are public, and all of the code used for this project is available
publicly under an open-source license, power users can generate any of the figures in this
paper or any other data outputs straightforwardly.
